---
output: 
  github_document:
    keep_html: true
#pdf_document
always_allow_html: true

---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# genRCT

### Generalizing the average treatment effect (ATE) from the trial using observational studies (OS)

### Installation
```{r, eval = FALSE}
devtools::install_github("idasomm/genRCT", 
                         auth_token = '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')

```



### Usage

genRCT(Y.trial, X.trial, A.trial, Y.rwe, X.rwe, A.rwe, family = "gaussian", 
       estimators = c("Naive", "IPSW", "AIPSW", "CW", "ACW-t", "ACW-b"), sieve = TRUE,
       inference = TRUE, n.boot = 100, conf.level = 0.05, seed = NULL, 
       plot.boot = TRUE, verbose = TRUE)

### Help functions
```{r, eval = FALSE}
library(genRCT)
?genRCT
```

### Arguments

Argument  | 
------------- | -------------
Y.trial | Observed outcome from a trial; vector of size $n$ (the trial sample size).
A.trial | Treatment received from a trial; vector of size $n$.
X.trial | Matrix of $p$ baseline covariates from a trial; dimension $n$ by $p$.
Y.rwe | Observed outcome from OS; if obtained, vector of size $m$ (OS sample size); otherwise, set Y.rwe = NULL.
A.rwe | Treatment received from OS; if obtained, vector of size $m$; otherwise, set Y.rwe = NULL.
X.rwe | Matrix of $p$ baseline covariates from OS; dimension $m by $p$.
family | The type of outcome; "gaussian" for continuous outcome or "binomial" for binary outcome. Default is "gaussian".
estimators | A vector of one or multiple methods to estimate the ATE. Allowed values are "Naive", "IPSW", "AIPSW", "CW", "ACW-t", "ACW-b". The "ACW-b" is allowed only when both "Y.rwe" and "A.rwe" are obtained. Default specifies all 6 methods.
sieve | A logical value indicating whether the method of sieves are used for estimating sampling score and outcome models. Used only if estimators = "AIPSW or "ACW-t" or "ACW-b". Default is TRUE.
inference | A logical value indicating whether inference for the ATE via bootstrap should be provided. Default it TRUE.
n.boot | A numeric value indicating the number of bootstrap samples used. This is only relevant if inference = TRUE. Default is 100.
conf.level | The level of bootstrap confidence interval; Default is 0.05.
seed | An optional integer specifying an initial randomization seed for reproducibility. Default is NULL, corresponding to no seed.
plot.boot | A logical value indicating whether histograms of the bootstrap samples should be produced. Default is TRUE.
verbose | A logical value indicating whether intermediate progress messages should be printed. Default is TRUE.


### Value

Value  | 
------------- | -------------
fit	|  A table of estimated ATEs with bootstrap SE and confidence interval.
plot | A set of histograms displaying the distribution of the bootstrapped estimates. The red vertical reference lines represent the estimated ATEs from each method.

### Example

```{r}

library(genRCT)
library(tidyverse)
library(data.table)

data("simulData")

data.trial <- simulData %>% dplyr::filter(delta == 1)
data.rwe <- simulData %>% dplyr::filter(delta == 0)

Y.trial <- as.vector(data.trial %>% dplyr::select(Y))
A.trial <- as.vector(data.trial %>% dplyr::select(A))
X.trial <- data.trial %>% dplyr::select(X1:X5)

Y.rwe <- as.vector(data.rwe %>% dplyr::select(Y))
A.rwe <- as.vector(data.rwe %>% dplyr::select(A))
X.rwe <- data.rwe %>% dplyr::select(X1:X5)

fit <- genRCT(Y.trial = Y.trial, A.trial = A.trial, X.trial = X.trial, Y.rwe = Y.rwe, A.rwe = A.rwe, X.rwe = X.rwe, family = "gaussian",
              estimators = c("Naive", "IPSW", "AIPSW", "CW", "ACW-t", "ACW-b"), sieve = TRUE,
              inference = TRUE, n.boot = 500, conf.level = 0.05, seed = 12345, plot.boot = TRUE, verbose = FALSE)

fit$fit

truth <- 27.4
fit$plot + geom_vline(xintercept = truth, color = 'blue', linetype = "dashed", show.legend = T)
```
The red solid lines represent the estimated ATEs and the blue dashed lines represent the true ATE.
The differences are

```{r echo = F}
truth - fit$fit[ ,1]
```





