---
output: 
  github_document:
    keep_html: true
#pdf_document
always_allow_html: true

---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# genRCT

### Generalizing the average treatment effect (ATE) from the trial using observational studies (OS)

### Installation
```{r, eval = FALSE}
devtools::install_github("idasomm/genRCT")
```



### Usage

genRCT(Y.trial, X.trial, A.trial, Y.rwe, X.rwe, A.rwe, family = "gaussian", 
       estimators = c("Naive", "IPSW", "AIPSW", "CW", "ACW-t", "ACW-b"), sieve = TRUE,
       inference = TRUE, n.boot = 100, conf.level = 0.05, seed = NULL, 
       plot.boot = TRUE, verbose = TRUE)

### Help functions
```{r, eval = FALSE}
library(genRCT)
?genRCT
```

### Arguments

Argument  | 
------------- | -------------
Y.trial | Observed outcome from a trial; vector of size n (trial sample size).
A.trial | Treatment received from a trial; vector of size n.
X.trial | Matrix of p baseline covariates from a trial; dimension n by p.
Y.rwe | Observed outcome from OS; if obtained, vector of size m (OS sample size); otherwise, set Y.rwe = NULL.
A.rwe | Treatment received from OS; if obtained, vector of size m; otherwise, set A.rwe = NULL.
X.rwe | Matrix of p baseline covariates from OS; dimension m by p.
family | The type of outcome; "gaussian" for gaussian regression or "binomial" for logistic regression Default is "gaussian".
estimators | A vector of one or multiple methods to estimate the ATE. Allowed values are "Naive", "IPSW", "AIPSW", "CW", "ACW-t", "ACW-b". The "ACW-b" is allowed only when both "Y.rwe" and "A.rwe" are obtained. Default specifies all 6 methods.
sieve | A logical value indicating whether the method of sieves are used for estimating sampling score and outcome models. Used only if estimators = "AIPSW" or "ACW-t" or "ACW-b". Default is TRUE.
inference | A logical value indicating whether inference for the ATE via bootstrap should be provided. Default it TRUE.
n.boot | A numeric value indicating the number of bootstrap samples used. This is only relevant if inference = TRUE. Default is 100.
conf.level | The level of bootstrap confidence interval; Default is 0.05.
seed | An optional integer specifying an initial randomization seed for reproducibility. Default is NULL, corresponding to no seed.
plot.boot | A logical value indicating whether histograms of the bootstrap samples should be produced. Default is TRUE.
verbose | A logical value indicating whether intermediate progress messages should be printed. Default is TRUE.


### Value

Value  | 
------------- | -------------
fit	|  A table of estimated ATEs with bootstrap SE and confidence interval.
plot | A set of histograms displaying the distribution of the bootstrapped estimates. The red vertical reference lines represent the estimated ATEs from each method.

### Example - gaussian

```{r}

library(genRCT)
library(tidyverse)
library(data.table)

data("simulData")

data.trial <- simulData %>% dplyr::filter(delta == 1)
data.rwe <- simulData %>% dplyr::filter(delta == 0)

Y.trial <- data.trial %>% dplyr::select(Y)
A.trial <- data.trial %>% dplyr::select(A)
X.trial <- data.trial %>% dplyr::select(X1:X5)

Y.rwe <- data.rwe %>% dplyr::select(Y)
A.rwe <- data.rwe %>% dplyr::select(A)
X.rwe <- data.rwe %>% dplyr::select(X1:X5)

fit <- genRCT(Y.trial = Y.trial, A.trial = A.trial, X.trial = X.trial, Y.rwe = Y.rwe, 
              A.rwe = A.rwe, X.rwe = X.rwe, family = "gaussian",
              estimators = c("Naive", "IPSW", "AIPSW", "CW", "ACW-t", "ACW-b"), 
              sieve = TRUE, inference = TRUE, n.boot = 500, conf.level = 0.05, 
              seed = 12345, plot.boot = TRUE, verbose = FALSE)

fit$fit

truth <- 27.4
fit$plot + geom_vline(xintercept = truth, color = 'blue', linetype = "dashed")
```
The red solid lines represent the estimated ATEs and the blue dashed lines represent the true ATE.
The differences are

```{r echo = F}
truth - fit$fit[ ,1]
```


### Example - survival
```{r}

data("simulData.surv")
library(tidyverse)


data.trial <- simulData.surv$trial
Y.trial <- data.trial %>% dplyr::select(Y)
d.trial <- data.trial %>% dplyr::select(d)
A.trial <- data.trial %>% dplyr::select(A)
X.trial <- data.trial %>% dplyr::select(X1:X3)

X.rwe <- simulData.surv$rwe


tau <- seq(10, 50, by = 10)
tau <- round(tau)



library(pracma)
library(ggplot2)
library(survival)
library(data.table)
library(nleqslv)
library(ncvreg)
library(MASS)
library(Matrix)


## Fit genRCT survival
fit <- genRCT.surv(Y.trial = Y.trial, d.trial = d.trial, A.trial = A.trial, X.trial = X.trial, X.rwe = X.rwe, 
                   tau = tau, n.boot = 100, conf.level = 0.05, seed = 123, verbose = FALSE)
```
```{r echo = F}

lower1 <- fit$surv$lower1
lower0 <- fit$surv$lower0
upper1 <- fit$surv$upper1
upper0 <- fit$surv$upper0

sv1 <- as.data.frame(fit$surv$surv1) %>% mutate(est = 's1', time = fit$surv$time1)
sv0 <- as.data.frame(fit$surv$surv0) %>% mutate(est = 's0', time = fit$surv$time0)
surv.comb <- rbind(sv1, sv0)

datas <- surv.comb %>% pivot_longer(cols = c(surv.nv:denom.dacw.sv), names_to = "method", values_to = "surv") %>%
    mutate(estimators = factor(method, levels = c("denom.nv", "surv.nv", "denom.dr", "surv.dr", "denom.or", "surv.or", "denom.ipsw", "surv.ipsw", 
                                                  "denom.cw", "surv.cw", "denom.dacw", "surv.dacw", "denom.dacw.sv", "surv.dacw.sv"),
                             labels = c("Naive", "Naive2", "DR", "DR2", "OR", "OR2", "IPSW", "IPSW2", "CW", "CW2", "ACW1", "ACW2", "ACW1(S)", "ACW2(S)")))

lower <- data.frame(rbind(lower1, lower0)) %>% pivot_longer(cols = c(surv.nv:denom.dacw.sv), names_to = "method", values_to = "lower") %>%
  mutate(estimators = factor(method, levels = c("denom.nv", "surv.nv", "denom.dr", "surv.dr", "denom.or", "surv.or", "denom.ipsw", "surv.ipsw", 
                                                "denom.cw", "surv.cw", "denom.dacw", "surv.dacw", "denom.dacw.sv", "surv.dacw.sv"),
                             labels = c("Naive", "Naive2", "DR", "DR2", "OR", "OR2", "IPSW", "IPSW2", "CW", "CW2", "ACW1", "ACW2", "ACW1(S)", "ACW2(S)")))

upper <- data.frame(rbind(upper1, upper0)) %>% pivot_longer(cols = c(surv.nv:denom.dacw.sv), names_to = "method", values_to = "upper") %>%
  mutate(estimators = factor(method, levels = c("denom.nv", "surv.nv", "denom.dr", "surv.dr", "denom.or", "surv.or", "denom.ipsw", "surv.ipsw",
                                                "denom.cw", "surv.cw", "denom.dacw", "surv.dacw", "denom.dacw.sv", "surv.dacw.sv"),
                             labels = c("Naive", "Naive2", "DR", "DR2", "OR", "OR2", "IPSW", "IPSW2", "CW", "CW2", "ACW1", "ACW2", "ACW1(S)", "ACW2(S)")))


datas$lower <- lower$lower
datas$upper <- upper$upper

lower1.w <- apply(fit$surv$surv1 - 1.96 * fit$surv$se1, 2, mono.dec)
upper1.w <- apply(fit$surv$surv1 + 1.96 * fit$surv$se1, 2, mono.dec)
lower0.w <- apply(fit$surv$surv0 - 1.96 * fit$surv$se0, 2, mono.dec)
upper0.w <- apply(fit$surv$surv0 + 1.96 * fit$surv$se0, 2, mono.dec)

lower.w <- data.frame(rbind(lower1.w, lower0.w)) %>% pivot_longer(cols = c(surv.nv:denom.dacw.sv), names_to = "method", values_to = "lower") %>%
  mutate(estimators = factor(method, levels = c("denom.nv", "surv.nv", "denom.dr", "surv.dr", "denom.or", "surv.or", "denom.ipsw", "surv.ipsw", 
                                                "denom.cw", "surv.cw", "denom.dacw", "surv.dacw", "denom.dacw.sv", "surv.dacw.sv"),
                             labels = c("Naive", "Naive2", "DR", "DR2", "OR", "OR2", "IPSW", "IPSW2", "CW", "CW2", "ACW1", "ACW2", "ACW1(S)", "ACW2(S)")))
upper.w <- data.frame(rbind(upper1.w, upper0.w)) %>% pivot_longer(cols = c(surv.nv:denom.dacw.sv), names_to = "method", values_to = "upper") %>%
  mutate(estimators = factor(method, levels = c("denom.nv", "surv.nv", "denom.dr", "surv.dr", "denom.or", "surv.or", "denom.ipsw", "surv.ipsw",
                                                "denom.cw", "surv.cw", "denom.dacw", "surv.dacw", "denom.dacw.sv", "surv.dacw.sv"),
                             labels = c("Naive", "Naive2", "DR", "DR2", "OR", "OR2", "IPSW", "IPSW2", "CW", "CW2", "ACW1", "ACW2", "ACW1(S)", "ACW2(S)")))
datas$lower <- lower.w$lower
datas$upper <- upper.w$upper

datas %>% filter(estimators %in% c("Naive", "DR", "OR", "IPSW", "CW", "ACW1", "ACW2", "ACW1(S)", "ACW2(S)")) %>%
  #filter(estimators %in% c("Naive2", "DR2", "OR2", "IPSW2", "CW2", "ACW", "ACW2")) %>%
  ggplot(aes(x = time)) + geom_line(aes(y = surv, color = est)) +
  # geom_rect(data = fake.ribbon,
  #           aes(xmin=time, xmax=time.max, ymin=lower2,
  #               ymax=upper2, fill=est),
  #           alpha=0.3, colour=NA) +
  geom_ribbon(aes(ymin=lower, ymax=upper, fill=est),
              alpha=0.3, colour=NA) +
  #geom_line(aes(y = lower, color = est), linetype = "dashed") +
  facet_wrap( ~ estimators, scales = "free", nrow=2) +
  coord_cartesian(xlim = c(0, 30)) +
  theme_bw()

```


